Properties of winning Iterated Prisoner's Dilemma strategies.

### 2020-03-30 paper/main.tex
Revise conclusion

This manuscript has explored the performance of \numberofstrategies strategies
of the Iterated Prisoner's Dilemma in a large number of computer tournaments.
The \numberofstrategies strategies used in this manuscript have been mainly for
the literature, accessible due to an open source software library
called the Axelrod-Python library. The software was used to generate a total of
\numberofalltournaments computer tournaments results with different number of
strategies and different participants each time. The data collection was
described in Section~\ref{section:data_collection}.

In
Section~\ref{section:top_performances}, the tournaments results were used to
present the top performances. The data set contained results from four different
settings, and these were also studied individually. In standard tournaments
complex strategies trained using reinforcement learning ranked in the top spots.
Some of these strategies ranked again in the top spots in probabilistic
ending tournaments when a \(p_e\) of less 0.1 was considered and in noisy tournaments
when \(p_n\) was less than 0.1. In probabilistic
ending tournaments \(p_e\) was designed to vary between 0 and 1. It was demonstrated
that for values larger than 0.1, as stated in the Folk Theorem, defecting strategies
were winning the tournaments because there was a high likelihood of the game
ending in the next turn. In tournaments with noise the median ranks of the top
15 strategies had the highest values and the \(r\) distributions were bimodal.
The top rank strategies were performing both well and bad, and this indicates
that in noisy tournaments there are not strategies that can guarantee winning.
However, if the probability of noise was constrained at 0.1 then strategies
designed for noisy tournaments indeed performed well.

Section~\ref{section:evaluation_of_performance} covered an analysis of
performance based on several features associated with a strategy and with the
environments in which it was competing. The results of this analysis showed that a
strategy's characteristics such as whether or not it's stochastic, and the information it
used regarding the game had no effect on the strategy's success. The most
important features have been those that compared the strategy's behaviour to its
environment. The cooperating ratio of the strategy compared to the mean
cooperator was highlighted as the most important feature in the analysis.
Other significant features included the SSE of a strategy and it's cooperating
rate after being in either state \(CC\) and \(CD\).

So what is the best way of playing the game? and is there a single dominant
strategy in the IPD? There is not a single strategy, from the collection of the
195 strategies, that has managed to perform well in all the tournaments it competed in.
Even if on average a strategy ranked highly in a specific environment it did not
guarantee its success over the different tournament types. However, the results
of sections~\ref{section:top_performances}
and~\ref{section:evaluation_of_performance} have demonstrated that there are
properties associated with the success of strategies. A few the properties that
have been identified by this manuscript's analysis contradict the properties of
Axelrod~\cite{Axelrod1981}. Namely, in Section~\ref{section:top_performances} it
was shown that trained strategies and strategies that decided their actions
based on pre-designed strategies to maximise their utility dominated several
tournaments across tournament types, hinting that successful IPD strategies are
often clever or more complex than simple strategies like TFT. Most of the
successful strategies highlighted
in~Section~\ref{section:top_performances} were strategies that begin with
cooperation.

@@

Thus, the five properties successful strategies need to have in a IPD competition
are: be nice, be provocable and contrite, be a little envious, be clever, and adapt to the
environment (the population of strategies).

>>>>

This manuscript has explored the performance of \numberofstrategies strategies
of the Iterated Prisoner's Dilemma in a large number of computer tournaments.
We analyzed and extracted the salient features of the best performing strategies
across various tournament types, casting the results in terms of Axelrod's
original suggested features of good IPD strategies. Moreover, our results
shed light on the historic performance of TFT, zero determinant strategies,
and memory-one strategies generally. Strategies need to match their play
to the cooperativeness of the tournament population and do so in a way that
prevents or minimizes exploitation. Overall we see that complex or clever
strategies can be effective, whether trained against a corpus of possible
opponents or purposely designed to mitigate the impact of noise such as
in the strategy DBS. Further, we showed that while the type of exploitation
attempted by ZD strategies is not typically effective in tournaments, more
sophisticated strategies capable of selectively exploiting weaker opponents
while mutually cooperating with stronger opponents can be highly successful.
This fact was also indicated numerically by the importance of the strategy
feature SSE in the analysis of strategy features,

Highly noisy or tournaments with short matches favor less cooperative strategies.
These environments mitigate the value of being nice. The
added uncertainty reduces the ability of maintaining mutual cooperation and
also triggers grudging strategies. Accordingly we find that in noisy tournaments
that the best performing players cooperate a lower rate than the tournament
population on average. Nevertheless we found some strategies designed or trained
for noisy environments were also highly ranked in noise-free tournaments. This
indicates that strategy complexity is not necessarily a liability, rather it
can confer adaptability to a more diverse set of environments.

In Section~\ref{section:top_performances}, the tournaments results were used to
present the top performances. The data set contained results from four different
settings, and these were also studied individually. In standard tournaments
complex strategies trained using reinforcement learning ranked in the top spots.
Some of these strategies ranked again in the top spots in probabilistic
ending tournaments when a \(p_e\) of less 0.1 was considered and in noisy tournaments
when \(p_n\) was less than 0.1. In probabilistic
ending tournaments \(p_e\) was designed to vary between 0 and 1. It was demonstrated
that for values larger than 0.1, as stated in the Folk Theorem, defecting strategies
were winning the tournaments because there was a high likelihood of the game
ending in the next turn. In tournaments with noise the median ranks of the top
15 strategies had the highest values and the \(r\) distributions were bimodal.
The top rank strategies were performing both well and bad, and this indicates
that in noisy tournaments where the noise can vary substantially,
there were no strategies that can guarantee winning across a range of noise.
However, if the probability of noise was constrained at 0.1 then strategies
designed for noisy tournaments indeed performed well.

So what is the best way of playing the IPD? And is there a single dominant
strategy for the IPD? There was not a single strategy within the collection of the
\numberofstrategies strategies, that has managed to perform well in all the
tournaments variations it competed in.
Even if on average a strategy ranked highly in a specific environment it did not
guarantee its success over the different tournament types. However, the results
of sections~\ref{section:top_performances}
and~\ref{section:evaluation_of_performance} have demonstrated that there are
properties associated with the success of strategies. A few the properties that
have been identified by this manuscript's analysis contradict the properties of
Axelrod~\cite{Axelrod1981}. Namely, in Section~\ref{section:top_performances} it
was shown that trained strategies and strategies that decided their actions
based on pre-designed strategies to maximise their utility dominated several
tournaments across tournament types, hinting that successful IPD strategies are
often clever or more complex than simple strategies like TFT. Most of the
successful strategies highlighted
in~Section~\ref{section:top_performances} were strategies that begin with
cooperation.

@@

Thus, the five properties successful strategies need to have in a IPD competition
are: be nice, be provocable and contrite, be a little envious, be clever, and adapt to the
environment (including the population of strategies).
